---
title: "SDS 322E: Project 1 Report"
output: html_document
---

## Names and EIDs of Group Members
Stephanie Kuo sk59344
Tony Huang tkh699
Annika Shivam aes5956



## Load the Data

```{r setup, include=FALSE}
## Do not modify this code chunk
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```



```{r, cache=TRUE, message=FALSE}
## This chunk may take a long time to run. Set 'cache=TRUE' in the chunk options
## to make it run faster in subsequent knittings.

championships <- read_tsv("WCA_export_championships.tsv.bz2")
competitions <- read_tsv("WCA_export_Competitions.tsv.bz2")
continents <- read_tsv("WCA_export_Continents.tsv.bz2")
countries <- read_tsv("WCA_export_Countries.tsv.bz2")
eligible_country_iso2s_for_championship <- read_tsv("WCA_export_eligible_country_iso2s_for_championship.tsv.bz2")
events <- read_tsv("WCA_export_Events.tsv.bz2")
formats <- read_tsv("WCA_export_Formats.tsv.bz2")
persons <- read_tsv("WCA_export_Persons.tsv.bz2")
ranksaverage <- read_tsv("WCA_export_RanksAverage_333.tsv.bz2")
rankssingle <- read_tsv("WCA_export_RanksSingle_333.tsv.bz2")
results <- read_tsv("WCA_export_Results_333.tsv.bz2")
roundtypes <- read_tsv("WCA_export_RoundTypes.tsv.bz2")
scrambles <- read_tsv("WCA_export_Scrambles.tsv.bz2")
```


## Questions

### Active Speed Cubers

How many active (3x3x3) speedcubers are there registered with the WCA? For this question an *active speeedcuber* is defined as any person registered in the WCA who has competed in at least two competitions in the years 2022--2024. 


```{r}
# filter competitions between 2022 and 2024
comp_22_24 <- competitions|>
  filter(year >= 2022 & year <= 2024)

# save all active cubers in a new dataset and count how many there are
active_cuber <- results|>
  #join on competitionId to filter 22-24 competitions
  inner_join(comp_22_24, by = c("competitionId" = "id"))|>
  group_by(personId)|>
  summarise(num_competition = n_distinct(competitionId))|>
  filter(num_competition >= 2)
nrow(active_cuber)

```

**There are 39482 active 3x3x3 speedcubers that are registered with th WCS between 2022-2024, measured by if they have competed in at least 2 competitions.**

### World Records

This question has two parts:

1. Who holds the current world record single? 

```{r}
#This code looks for the current world record single by joining different datasets
curr_world_record <- rankssingle|>
  filter(worldRank == 1)|>
  # join to get name of world Rank 1 people
  inner_join(persons, by = c("personId" = "id"))|>
  # join with result to find competition
  inner_join(results, by = c("personId", "best"))|>
  select(personName, best, competitionId, worldRank)|>
  slice(1)
curr_world_record
```

**Max Park currently hold the world record single.**


On what date was this record set?

NOTE: Use the start date of the competition in which the record was set.

```{r}
## This builds off of the previous code by joining with competition to get start date
curr_world_record <- rankssingle|>
  filter(worldRank == 1)|>
  # join to get name of world Rank 1 people
  inner_join(persons, by = c("personId" = "id"))|>
  # join with result to find competition
  inner_join(results, by = c("personId", "best"))|>
  # join with competition to get the date
  inner_join(competitions, by = c("competitionId" = "id"))|>
  select(personName, best, competitionId, worldRank, year, month, day)|>
  slice(1)
curr_world_record


```

**On June 11, 2023, the world record was set by Max Park.**



2. Who *previously* held the world record single? On what date was this previous record set?

```{r}
## Find who previously held the world record single and the date the record was set
prev_world_record <- results|>
  filter(regionalSingleRecord == "WR")|>
  # join with competition to get the date
  inner_join(competitions, by = c("competitionId" = "id"))|>
  # joining with persons to get name
  inner_join(persons, by = c("personId" = "id"))|>
  select(personName, best, competitionId, year, month, day)|>
  arrange(best)|>
  slice(2)
prev_world_record


```

**Yusheng Du (杜宇生) previously held the world record single on November 24, 2018.**


### Regional Rankings

This question has two parts:

1. Amongst all speedcubers, who is the top ranked male speedcuber (for best single solve) in Australia?

```{r}
## Add your code here

```

**Write your answer here.**


2. Amongst all speedcubers, who is the top ranked female speedcuber (for best single solve time) in Europe?

```{r}
## Add your code here

```

**Write your answer here.**



### Time Until Sub-5

Having a time below 5 seconds is considered an elite achievement and most speedcubers have to complete a large number of solves before they can obtain a sub-5 second solve. 

**NOTE**: Each round of a competition has 5 solves that should be considered separately when counting the number of solves.


1. For the current top 10 speedcubers in the world (as recorded in the RanksSingle table), on average, how many completed solves did they have to do before achieving a sub-5 second solve?


```{r}
## Add your code here

```

**Write your answer here.**



2. For **one** of the top 10 speedcubers make a plot of their solve times vs. the date of the solve, with date on the x-axis and solve time on the y-axis. 


```{r}
## Add your code here

```


**Write your answer here.**


### Up-and-Coming Speed Cubers

Which speed cubers **not** in the top 10,000 (worldwide for single best time) should we keep an eye on for the near future? 

The idea here is to identify "up-and-coming" speedcubers who are not yet achieving elite times. Come up with a list of **five** speedcubers (provide their names and WCA IDs) that you have identified as "up-and-coming". There is no one way to answer this question and the goal is to provide an analysis of the data that justifies the selection of your five names.

```{r}
## Add your code here


```

**Write your answer here.**



### Region Rivalries

Europe and North America are both regions with strong speedcubers in the WCA. 

Which region has the faster group of speedcubers on average? 

To answer this question, characterize each person using their best *average* score according to their listing in the `ranksaverage` table. In the `persons` table the `countryId` indicates each person's country affiliation. The `countries` table lists the region that each country is in via the `continentId` column (Europe is "_Europe" and North America is "_North America").

Before attempting to answer the question, state what you expect the answer to be below.

**Write your expectation here.**



```{r}
## Add your code here

```

What do you conclude about speedcubers in Europe vs. North America?

**Write you answer here.**


### Alternative Explanations

Develop an alternative explanation/hypothesis regarding speedcubers from Europe and North America that is 

1. Consistent with the results you produced in the previous question; but

2. Provides a different interpretation or explanation for what is going on. 

If the results from the previous question were unexpected, make use of systems thinking to develop an alternative hypothesis. If the results were consistent with your expectations, then use skeptical thinking. In either case, you should present an analysis that shows evidence for or against this alternative explanation relative to the conclusion that you made in the previous question.

**State your alternative explanation here.**

```{r}
## Add your code here

```

**Summarize the results of your alternative analysis here.**





## Discussion
Putting it all together, what did you learn from your data?
* Reflect on the process of conducting this project. What was challenging, what have you
learned from the process itself?

* Was there anything unexpected that you found in your analysis? Or was everything
essentially as you expected?

* Include acknowledgements for any help received

* Report the contribution of each member (i.e. who did what).

**answer goes here**
